Robust Trajectory Optimization on a Quantum System

### 2021-02-05 paper/6stochastic.tex
jens edits 2 addressed or todo'd

\section{Robustness to Time-Dependent Parameter Uncertainty \label{sec:stochastic}}
An additional source of experimental error arises from time-dependent
Hamiltonian parameter uncertainty. For many flux-biased and inductively-coupled
superconducting circuit elements, magnetic flux noise is a significant
source of coherent errors. Flux noise
modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
by $a(t) \rightarrow a(t) + \delta a(t)$.
The spectral density of the flux noise $\delta a(t)$ is observed to
follow a 1/$f$ distribution
\cite{bialczak20071f, koch2007model,
  yoshihara2006decoherence, yoshihara2010correlated},
so the noise is dominated by low-frequency components.
The analytic gate considered here
takes advantage of the low-frequency characteristic and
treats the noise as quasi-static, performing a generalization of the spin-echo
technique to compensate for erroneous drift \cite{hahn1952spin, meiboom1958modified}.

>>>>

\section{Robustness to Time-Dependent Parameter Uncertainty \label{sec:stochastic}}
An additional source of experimental error arises from time-dependent
parameter uncertainty. For many flux-biased and inductively-coupled
superconducting circuit elements, magnetic flux noise is a significant
source of coherent errors. Flux noise
modifies the fluxonium Hamiltonian \eqref{eq:hamiltonian}
by $a(t) \rightarrow a(t) + \delta a(t)$.
The spectral density of the flux noise $\delta a(t)$ is observed to
follow a 1/$f$ distribution
\cite{bialczak20071f, koch2007model,
  yoshihara2006decoherence, yoshihara2010correlated},
so the noise is dominated by low-frequency components.
The analytic gate considered here
takes advantage of the low-frequency characteristic and
treats the noise as quasi-static, performing a generalization of the spin-echo
technique to compensate for erroneous drift \cite{hahn1952spin, meiboom1958modified}.

### 2021-02-05 paper/a1longitude.tex
jens edits 2 addressed or todo'd

\section{Depolarization \label{appendix:longitude}}
We comment on the depolarization metrics and then give
our procedure for integrating the Lindblad master equation.
The depolarization probability and the gate error due to
depolarization metrics are compared in Table \ref{tab:longitude} for the
numerical experiment described in Section \ref{sec:longitude}.
The relative performance of the analytic and numerical techniques
is similar across the two metrics.

\begin{table}[ht]
  \begin{tabular}{c | c | c | c | c | c | c}
         & $P_{1\textrm{A}}$ & $P_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
    Gate & $(10^{-5})$ & $(10^{-5})$ & $P_{1\textrm{A}}/ P_{1\textrm{N}}$ & $(10^{-5})$ &
    $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
    Y/2 & 5.253 & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
  \end{tabular}
  \caption{
    Single gate depolarization probability (P)
    and single gate error due to depolarization (GE).
    Values are reported for the analytic (A) and numerical (N) gates.
  }
  \label{tab:longitude}
\end{table}

We employ the Lindblad master equation
to compute the gate error due to depolarization.
This equation takes the form:
\begin{equation}
  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
  + \sum_{i = 1}^{n^{2} - 1} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
\end{equation}
where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix, $n$ is the
dimension of the Hilbert space,
$[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
anti-commutator.
For depolarization $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
$\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,
$L_{\uparrow} = \sigma^{+}/2$, and
$L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained from the spline
shown in Figure \ref{fig:longitudeb}.
We obtain the $T_{1}$ values in this spline
by driving the qubit at the desired flux amplitude
and monitoring the resultant decay. For more details
on these measurements, consult \cite{zhang2020universal}.

So that we may use exponential integrators, we employ
the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
\begin{equation}
  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho}),
\end{equation}
\begin{equation}
  \begin{aligned}
    \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
    &+ \sum_{i = 1}^{n^{2} - 1} \gamma_{i}
    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
    - L_{i}^{T}L_{i}^{*} \otimes I)),
  \end{aligned}
\end{equation}
where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
We use zero-order hold on the controls so the exact solution is
$\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k} \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
This isomorphism transforms $(n^{2} \times n^{2}) \times (n^{2} \times n^{2})$
matrix-matrix multiplications to $(n^{4} \times n^{4}) \times n^{4}$ matrix-vector
multiplications. For small $n$ and zero-order hold on the controls, we find that it is
faster to use an exponential integrator on the vectorized equation than to perform
Runge-Kutta on the unvectorized equation. The latter requires decreasing the integration
time step to maintain accuracy, resulting in more knot points.

>>>>

\section{Depolarization \label{appendix:longitude}}
We comment on the depolarization metrics and then give
our procedure for integrating the Lindblad master equation.
The integrated depolarization rate and the gate error due to
depolarization metrics are compared in Table \ref{tab:longitude} for the
numerical experiment described in Section \ref{sec:longitude}.
The relative performance of the analytic and numerical techniques
is similar across the two metrics.

\begin{table}[ht]
  \begin{tabular}{c | c | c | c | c | c | c}
         & $D_{1\textrm{A}}$ & $D_{1\textrm{N}}$ & & GE$_{\textrm{A}}$ & GE$_{\textrm{N}}$ &\\
    Gate & $(10^{-5})$ & $(10^{-5})$ & $D_{1\textrm{A}}/ D_{1\textrm{N}}$ & $(10^{-5})$ &
    $(10^{-5})$ & GE$_{\textrm{A}} / \textrm{GE}_{\textrm{N}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000 & 1.776 & 0.371 & 4.787\\
    Y/2 & 5.253 & 1.157 & 4.540 & 1.539 & 0.370 & 4.159\\
    X/2 & 16.251 & 2.660 & 6.109 & 5.347 & 0.863 & 6.196\\
  \end{tabular}
  \caption{
    Single gate integrated depolarization rate ($D_{1}$)
    and single gate error due to depolarization (GE).
    Values are reported for the analytic (A) and numerical (N) gates.
  }
  \label{tab:longitude}
\end{table}

We employ the Lindblad master equation
to compute the gate error due to depolarization.
This equation takes the form:
\begin{equation}
  \frac{d}{dt} \rho = \frac{-i}{\hbar} [H, \rho]
  + \sum_{i} \gamma_{i} (L_{i} \rho L_{i}^{\dagger}
  - \frac{1}{2} \{L_{i}^{\dagger} L_{i}, \rho\}),
\end{equation}
where $\rho = \ket{\psi}\bra{\psi}$ is the density matrix,
$[\cdot, \cdot]$ is the algebraic commutator, and $\{\cdot, \cdot \}$ is the algebraic
anti-commutator.
For depolarization $\gamma_{\uparrow} = T_{1, \uparrow}^{-1}$,
$\gamma_{\downarrow} = T_{1, \downarrow}^{-1}$,
$L_{\uparrow} = \sigma^{+}/2$, and
$L_{\downarrow} = \sigma^{-}/2$ where $\sigma^{\pm} = \sigma_{x} \pm i \sigma_{y}$.
Both $T_{1, \uparrow}$ and $T_{1, \downarrow}$ are obtained at each knot point
from the spline shown in Figure \ref{fig:longitudeb}.
We obtain the $T_{1}$ values in this spline
by driving the qubit at the desired flux amplitude
and monitoring the resultant decay. For more details
on these measurements, consult \cite{zhang2020universal}.

So that we may use exponential integrators, we employ
the Vectorization/Choi-Jamiolkowski isomorphism \cite{Landi2018},
\begin{equation}
  \frac{d}{dt} \textrm{vec}({\rho}) = \hat{\mathcal{L}} \textrm{vec}({\rho}),
\end{equation}
\begin{equation}
  \begin{aligned}
    \hat{\mathcal{L}} &= -i(I \otimes H - H^{T} \otimes I)\\
    &+ \sum_{i} \gamma_{i}
    (L_{i}^{*} \otimes L_{i} - \frac{1}{2} (I \otimes L_{i}^{\dagger}L_{i}
    - L_{i}^{T}L_{i}^{*} \otimes I)),
  \end{aligned}
\end{equation}
where $\rho = \sum_{i, j} \alpha_{i, j} \ket{i}\bra{j}$
and $\textrm{vec}(\rho) = \sum_{i, j} \alpha_{i, j} \ket{i} \otimes \ket{j}$.
We use zero-order hold on the controls--equivalently, $a(t)$ is piecewise-constant
on the intervals $[t_{k}, t_{k + 1}]$. Therefore,
$H(a(t))$, $L(a(t))$, and $\gamma_{i}(a(t))$ are also piecewise-constant.
So, the exact solution is
$\textrm{vec}(\rho_{k + 1}) = {\exp}{\textstyle(}\Delta t_{k} \hat{\mathcal{L}}_{k}{\textstyle)} \textrm{vec}(\rho_{k})$.
This isomorphism transforms $(n \times n) \times (n \times n)$
matrix-matrix multiplications to $(n^{2} \times n^{2}) \times n^{2}$ matrix-vector
multiplications. For small $n$ and zero-order hold on the controls, we find that it is
faster to use an exponential integrator on the vectorized equation than to perform
Runge-Kutta on the unvectorized equation. The latter requires decreasing the integration
time step to maintain accuracy, resulting in more knot points.

### 2021-02-05 paper/a2unscented.tex
jens edits 2 addressed or todo'd

\section{Unscented Transformation}
\label{appendix:unscented}
In this section, we outline the full unscented sampling procedure.
In the unscented sampling method, we consider a state $\psi \in \mathbb{R}^{2n}$, 
an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
The state is obtained with an isomorphism \eqref{eq:isomorphism} and
we omit bra-ket notation to emphasize that the state is real.
The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
positive-definite covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
which describes the uncertainty in the initial state.
$P_{1}$ is typically non-zero even if the state preparation error is negligible.
The uncertain parameter has zero-mean
and its distribution is given by the covariance matrix
$L_{k} \in \mathbb{S}_{++}^{d}$ at knot point $k$. The zero-mean assumption
is convenient for deriving the update procedure. A non-zero mean can be encoded
in the discrete dynamics function $f(\psi_{k}, \lambda_{k})$.

The initial $4n + 2d$ sample states and initial $4n + 2d$
uncertain parameters are sampled from the initial distributions,
\begin{equation}\label{eq:uupdate}
  \begin{bmatrix} \psi_{1}^{j} \\ \lambda_{1}^{j} \end{bmatrix} =
  \begin{bmatrix} \bar{\psi}_{1} \\ 0\end{bmatrix}
    \pm \beta \sqrt{\begin{bmatrix} P_{1} & 0\\ 0 & L_{1}\end{bmatrix}}^{\; j}.
\end{equation}
$\beta$ is a hyperparameter that controls the spacing of the covariance contour.
The $(\pm)$ is understood to take $(+)$ for $j \in \{1, \dots, 2n + d\}$ and $(-)$ for
$j \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
to compute the square root of the
joint covariance matrix, though other methods
such as the principal square root may be employed.
The superscript on the matrix square root indicates the $i$\textsuperscript{th}
column (mod $2n + d$) of the lower triangular Cholesky factor.
Then, the sample states are normalized,
\begin{equation}\label{eq:unormalize}
  \psi_{1}^{j} \rightarrow \frac{\psi_{1}^{j}}{\sqrt{{\psi_{1}^{j}}^{T} \psi_{1}^{j}}}.
\end{equation}
The sample states are propagated to the next knot point,
\begin{equation}\label{eq:upropagate}
  \psi^{j}_{2} = f(\psi^{j}_{1}, \lambda^{j}_{1}).
\end{equation}
The mean and covariance of the sample states are computed,
\begin{align}
  \bar{\psi}_{2} &= \frac{1}{4n + 2d} \sum_{j = 1}^{4n + 2d} \psi_{2}^{j},\\
  P_{2} &= \frac{1}{2 \beta^{2}} \sum_{j = 1}^{4n + 2d}
  (\psi^{j}_{2} - \bar{\psi}_{2})(\psi^{j}_{2} - \bar{\psi}_{2})^{T}.
\end{align}
The sample states are then resampled and propagated to the next knot point using
\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
choice of sample states (sigma points) follows equation 11 of
\cite{julier2004unscented}.
Prescriptions that require fewer sigma points exist \cite{julier2002reduced}.

>>>>

\section{Unscented Transformation}
\label{appendix:unscented}
In this section, we outline the full unscented sampling procedure.
In the unscented sampling method, we consider a state $\psi \in \mathbb{R}^{2n}$, 
an uncertain parameter $\lambda \in \mathbb{R}^{d}$, and discrete dynamics
$\psi_{k + 1} = f(\psi_{k}, \lambda_{k})$.
We omit bra-ket notation to emphasize that
the state is a real vector, and it is given by the right-hand-side of the
isomorphism \eqref{eq:isomorphism}.
The nominal initial state is given by $\bar{\psi}_{1}$ with an associated
covariance matrix $P_{1} \in \mathbb{S}_{++}^{2n}$
which describes the uncertainty in the initial state.
We use the notation $\mathbb{S}_{++}^{m}$ to denote the
set of real, symmetric, and positive-definite $m \times m$ matrices.
$P_{1}$ is typically non-zero even if the state preparation error is negligible.
The uncertain parameter has zero-mean
and its distribution is given by the covariance matrix
$L_{k} \in \mathbb{S}_{++}^{d}$ at knot point $k$. The zero-mean assumption
is convenient for deriving the update procedure. A non-zero mean can be encoded
in the discrete dynamics function $f(\psi_{k}, \lambda_{k})$.

The initial $4n + 2d$ sample states and initial $4n + 2d$
uncertain parameters are sampled from the initial distributions,
\begin{equation}\label{eq:uupdate}
  \begin{bmatrix} \psi_{1}^{j} \\ \lambda_{1}^{j} \end{bmatrix} =
  \begin{bmatrix} \bar{\psi}_{1} \\ 0\end{bmatrix}
    \pm \beta \sqrt{\begin{bmatrix} P_{1} & 0\\ 0 & L_{1}\end{bmatrix}}^{\; j}.
\end{equation}
$\beta$ is a hyperparameter that controls the spacing of the covariance contour.
The $(\pm)$ is understood to take $(+)$ for $j \in \{1, \dots, 2n + d\}$ and $(-)$ for
$j \in \{2n + d + 1, \dots, 4n + 2d\}$. We use the Cholesky factorization
to compute the square root of the
joint covariance matrix, though other methods
such as the principal square root may be employed.
The superscript on the matrix square root indicates the $j$\textsuperscript{th}
column (mod $2n + d$) of the lower triangular Cholesky factor.
Then, the sample states are normalized,
\begin{equation}\label{eq:unormalize}
  \psi_{1}^{j} \rightarrow \frac{\psi_{1}^{j}}{\sqrt{{\psi_{1}^{j}}^{T} \psi_{1}^{j}}}.
\end{equation}
The sample states are propagated to the next knot point,
\begin{equation}\label{eq:upropagate}
  \psi^{j}_{2} = f(\psi^{j}_{1}, \lambda^{j}_{1}).
\end{equation}
The mean and covariance of the sample states are computed,
\begin{align}
  \bar{\psi}_{2} &= \frac{1}{4n + 2d} \sum_{j = 1}^{4n + 2d} \psi_{2}^{j},\\
  P_{2} &= \frac{1}{2 \beta^{2}} \sum_{j = 1}^{4n + 2d}
  (\psi^{j}_{2} - \bar{\psi}_{2})(\psi^{j}_{2} - \bar{\psi}_{2})^{T}.
\end{align}
The sample states are then resampled and propagated to the next knot point using
\eqref{eq:uupdate}, \eqref{eq:unormalize}, and \eqref{eq:upropagate}. Our
choice of sample states (sigma points) follows equation 11 of
\cite{julier2004unscented}.
Prescriptions that require fewer sigma points exist \cite{julier2002reduced}.

### 2021-02-05 paper/a4time.tex
danny paper edits!
jens edits 2 addressed or todo'd

\section{Computational Performance \label{appendix:time}}
We provide runtimes for our optimizations and comment on the scaling of the
robustness methods. The runtimes for
the base optimization in Section \ref{sec:fluxonium},
the depolarization optimization in Section \ref{sec:longitude},
and the robust optimizations in Section \ref{sec:static}
are presented in Table \ref{tab:time}
for a $Z/2$ gate at gate times which are multiples of $1/4f_{q} \sim 18$ns.
We performed optimizations on a single core of
an AMD Ryzen Threadripper 3970X 32-Core Processor.
Future work will parallelize the robustness methods using GPUs
\cite{leung2017speedup},
which will enable fast optimizations on high-dimensional Hilbert spaces.

@@

Now we present the problem size complexities for the robustness methods.
For the sampling method, the size of the augmented state vector
is $O(dn^{3})$, where $d$ is the number of uncertain parameters and
$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
in the operator basis, $2d$ sample states per initial state,
and each state has $2n$ real numbers.
For the unscented sampling method, the size of the augmented state vector
is $O(dn^{3} + n^{4})$.
There are $n^{2}$ initial states in the operator basis,
$2(2n + d)$ sample states per initial state,
and each state has $2n$ real numbers.
For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
is $O(dmn^{3})$. There are $n^{2}$
initial states in the operator basis, $dm$ state derivatives per initial state,
and each state has $2n$ real numbers.

>>>>

\section{Computational Performance \label{appendix:time}}
We provide runtimes for our optimizations and comment on the problem size of the
robustness methods. The runtimes for
the base optimization in Section \ref{sec:fluxonium},
the depolarization optimization in Section \ref{sec:longitude},
and the robust optimizations in Section \ref{sec:static}
are presented in Table \ref{tab:time}
for a $Z/2$ gate at gate times which are multiples of $1/4f_{q} \sim 18$ns.
We performed optimizations on a single core of
an AMD Ryzen Threadripper 3970X 32-Core Processor.
Future work will parallelize the robustness methods using GPUs
\cite{leung2017speedup},
which will enable fast optimizations on high-dimensional Hilbert spaces.

@@

Now we note the size of the augmented state vector for the robustness methods.
For the sampling method, the size of the augmented state vector
is $O(dn^{3})$, where $d$ is the number of uncertain parameters and
$n$ is the dimension of the Hilbert space. There are $n^{2}$ initial states
in the operator basis, $2d$ sample states per initial state,
and each state has $2n$ real numbers.
For the unscented sampling method, the size of the augmented state vector
is $O(dn^{3} + n^{4})$.
There are $n^{2}$ initial states in the operator basis,
$2(2n + d)$ sample states per initial state,
and each state has $2n$ real numbers.
For the $m$\textsuperscript{th}-order derivative method, the size of the augmented state vector
is $O(dmn^{3})$. There are $n^{2}$
initial states in the operator basis, $dm$ state derivatives per initial state,
and each state has $2n$ real numbers. \todo{maybe add comment about auto-diff}
